# Visualisation

```{r echo=FALSE}
load("mod_results.RData")
load("mod2_results.RData")
load("final_reading_data.RData")
```

## The final reading dataset
A violin-box plot is created for the final reading dataset for analysis. 
```{r warning=FALSE}
ggplot(ReadingData_ENandNL, aes(x = V_Category, y = WORD_FIRST_FIXATION_DURATION, fill = V_Category)) +
  geom_violin(alpha = .6) +
  geom_boxplot(width = .2, alpha = .6) +
  stat_summary(fun = "mean", geom = "point",
               position = position_dodge(.9)) +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1,
               position = position_dodge(.9))+
  scale_x_discrete(name = "Valence Category") +
  scale_y_continuous(name = "Single Fixation Duration (ms)") +
  scale_fill_viridis_d(option = "E") + 
  facet_wrap(~Language) +
  theme_bw() +
  guides(fill = FALSE)
```

## The baseline model (mod)
I created a Blobbogram for Generalised linear mixed-effect model.broom.mixed::tidy() works on LMM to create a model summary. I then added factor to split the data for colours.
```{r message= FALSE, warning= FALSE}
modelsummary_vis <- broom.mixed::tidy(mod, effects = c("ran_pars", "fixed"),scales = NULL, ran_prefix = NULL, conf.int = TRUE, conf.level = 0.95, conf.method = "Wald") %>% 
  filter(effect =="fixed") %>% filter(term!="(Intercept)")

#Add factor to split the data for colours
modelsummary_vis$Type <- as.factor(ifelse(str_detect(modelsummary_vis$term, ":"), "Interaction", "Main effect"))
```

Below ggplot() coding was run to generate a Blobbogram. Condition1 refers to Positive vs Neutral; Condition2 refers to Negative vs Neutral. The minimal theme is applied to make it black & white = colourblind friendly.
```{r message= FALSE, warning= FALSE}
ggplot(modelsummary_vis, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high, shape = Type)) + 
  geom_pointrange() +
  ylab("Estimates with 95% CIs") + 
  geom_hline(aes(yintercept = 0), linetype = "dashed") +
  scale_x_discrete(limits = c("V_Category_NeuNeg_dev:Language_dev", "V_Category_NeuPos_dev:Language_dev", "WORD_LENGTH", "Conc_Mean",  "Language_dev", "V_Category_NeuNeg_dev", "V_Category_NeuPos_dev"),
                   labels = c("Condition2:Language", "Condition1:Language", "Word Length", "Concreteness", "Language", "Condition2", "Condition1")) +
  xlab("Fixed effect") +
  coord_flip() +
  theme_bw()
```

## emmip: Interaction-style plots for estimated marginal means
```{r message= FALSE, warning= FALSE}
plot_dat <- emmip(mod2, Language~V_Category, plotit = FALSE)

ggplot(plot_dat, aes(x = xvar, y = yvar, group = tvar, linetype = tvar, shape = tvar)) +
    geom_point(size = 2) +
    geom_line() +
    labs(x = "Valence Category", y = "Single Fixation Duration (ms)", linetype = "Language", shape = "Language") +
    theme_bw()
```

## Assumption Checking
1. Linearity
```{r message= FALSE, warning= FALSE}
plot(fitted(mod), residuals(mod))
```

2. Absence of collinearity


3. Homoskedasticity
Again, this can be checked with residual plots.
```{r message= FALSE, warning= FALSE}
plot(fitted(mod), residuals(mod))
```

4. Normality of residuals
```{r message= FALSE, warning= FALSE}
hist(residuals(mod))
qqnorm(residuals(mod))
```

5. Absence of influential data points
This assumption is met with visual investigation of violin boxplot.

6. Independence
This assumption is met with modeling (fixed effects: Valence cagetory, Language; random effects: Concreteness, Word Length). 

